### 2024-03-11 paper/header.tex
juliacon proceedings

% **************GENERATED FILE, DO NOT EDIT**************

\title{Fast, Elegant, Set-oriented Numerical Analysis using GAIO.jl}

\author[1]{A. Herwig}
\author[1]{O. Junge}
\affil[1]{Technical University Munich}

\keywords{Dynamical Systems}

\hypersetup{
pdftitle = {Fast, Elegant, Set-oriented Numerical Analysis using GAIO.jl},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {A. Herwig, O. Junge},
pdfkeywords = {Dynamical Systems},
}