### 2024-07-03 paper/header.tex
update paper plots

% **************GENERATED FILE, DO NOT EDIT**************

\title{GAIO.jl - a package for ...}

\author[2]{Michael Dellnitz}
\author[1]{April Herwig}
\author[1]{Oliver Junge}
\affil[1]{Technical University of Munich, Germany}
\affil[2]{University of Paderborn, Germany}

\keywords{dynamical system, invariant set, chain recurrent set, invariant manifold, attractor, invariant measure, almost invariant set, metastable set, coherent set, transfer operator, Koopman operator}

\hypersetup{
allcolors={blue},
pdftitle = {GAIO.jl - a concise package for the global analysis of dynamical systems},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {April Herwig, Oliver Junge},
pdfkeywords = {dynamical system, invariant set, chain recurrent set, invariant manifold, attractor, invariant measure, almost invariant set, metastable set, coherent set, transfer operator, Koopman operator},
}